---
title: 'Statisctical Inference Assignment Part-1: A simulation Exercise'
author: "Benhur Tedros"
date: "February 15, 2017"
output: html_document
---

## Instructions

The project consists of two parts:

#### 1. A simulation exercise.
#### 2. Basic inferential data analysis [this analysis is done in separate Rmd.file]


You will create a report to answer the questions. Given the nature of the series, ideally you'll use knitr to create the reports and convert to a pdf. 

### Part-1: Simulation Exercise

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

#### A) Set up the directory

```{r setup, include=TRUE}

setwd("F:/BENHUR FOLDER/Coursera/Statistical Reference/Week 4")
library(ggplot2)

```

#### B) List the given variables in the problem and define them

Number of average expontials: n=40
Lamda=0.2
Number of simulations: numsim=1000

```{r set up variables, echo=TRUE}
n<-40
lamda<-0.2
numsim<-1000

```

Set.seed() is essential when we generate randomly generated numbers such as simulation based estimation.
rexp () will provide the n*numsim from the exponential distribution with lamda as its rate parameter.
matrix () will rearrange the random variables (n*numsim).
Apply () will be used to calculate the mean of each row.
plots() will be generated for analysis.


```{r hist, echo=TRUE}

set.seed(13579)
subset<-matrix(rexp(n*numsim,rate=lamda), numsim)
subset_mean<-apply(subset,1,mean)
hist(subset_mean,xlab="Mean of the subset data", col="purple")

```

#### Answering the questions

##### A) To show the sample mean and compare it to the theoretical mean of the distribution.

```{r theoretical mean, echo=TRUE}

Theoretical_mean<-1/lamda
Theoretical_mean

```

Actual center of distribution based on simulation

```{r simulation mean, echo=TRUE}

round(mean(subset_mean),2)

```
##### B) To show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution

```{r theoretical variance and sd, echo=TRUE}

Theoretical_var<-(1/lamda)^2/n
Theoretical_var

Theoretical_sd<-(1/lamda)/sqrt(n)
Theoretical_sd

```
Actual variance and standard deviation based on simulation

```{r simulation variance and sd, echo=TRUE}

round(var(subset_mean),2)
round(sd(subset_mean),2)

```
##### C) To show that the distribution is approximately normal

  To prove the distribution of means are normal distributed (as CLT states), ggplot was used

```{r normal distribution, echo=TRUE}

data_normd<-data.frame(subset_mean)

g<-ggplot(data_normd,aes(x=subset_mean),xlab="distribution of means")

g +geom_histogram(aes(y=..density..),colour="blue",fill="orange")+geom_density() 

```

So, the result looks a normally distributed means.

